Replacing standard galaxy profiles with \\ mixtures of Gaussians

### 2012-10-22 documents/mixtures/mixture_models.tex
words about speed

% to-do
% -----
% - DSTN: we need *some* estimate of the speed-up this delivers to The Tractor in the abstract and in the text -- of course this is bullshit because we get to decide how precisely we were doing things prior to using the mixtures -- but we need it since it is our main point.
% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
% - DSTN: make figures showing use of the models in a real situation; see last figure
% - HOGG: does the Lupton paper give the profiles, or should we cite the code directly?
% - HOGG: need Bundy reference
% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
% - HOGG: send to Lupton and Bundy for comments
% - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
% - DSTN: is there anyone we need to ack on your side, or granting agencies?
% - HOGG: should we make Machine Readable Tables for PASP?

@@

\begin{abstract}
Exponential, de~Vaucouleurs, and S\'ersic profiles are simple and
successful models for fitting two-dimensional images of galaxies.  One
numerical issue encountered in this kind of fitting is the pixel
rendering and convolution (or correlation) of the models with the
telescope point-spread function (PSF); these operations are slow, and
easy to get slightly wrong at small radii.  Here we exploit the
realization that these models can be approximated to arbitrary
accuracy with a mixture (linear superposition) of two-dimensional
Gaussians (MoGs).  MoGs are fast to render and fast to
affine-transform.  Most importantly, if you have a MoG PSF model
(which we advocate), the PSF-convolved, affine-transformed galaxy
models are themselves MoGs and therefore very fast to compute,
integrate, and render precisely.  We present worked examples that can
be directly used in image fitting; we are using them ourselves.  The
MoG profiles we provide can be swapped in to replace the standard
models in any image-fitting code; they sped up model fitting in our
projects by XXX; they ought to make any code faster at very little
cost in accuracy.
\end{abstract}

@@

The speed-ups that can be obtained by using MoG approximations can be
very large... DSTN: How much did we benefit when we switched to MoGs
in \project{The Tractor}?...

>>>>

% to-do
% -----
% - DSTN: check sentences I say about ``ten times faster''.
% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
% - DSTN: make figures showing use of the models in a real situation; see last figure
% - HOGG: does the Lupton paper give the profiles, or should we cite the code directly?
% - HOGG: need Bundy reference
% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
% - HOGG: send to Lupton and Bundy for comments
% - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
% - DSTN: is there anyone we need to ack on your side, or granting agencies?
% - HOGG: should we make Machine Readable Tables for PASP?

@@

\begin{abstract}
Exponential, de~Vaucouleurs, and S\'ersic profiles are simple and
successful models for fitting two-dimensional images of galaxies.  One
numerical issue encountered in this kind of fitting is the pixel
rendering and convolution (or correlation) of the models with the
telescope point-spread function (PSF); these operations are slow, and
easy to get slightly wrong at small radii.  Here we exploit the
realization that these models can be approximated to arbitrary
accuracy with a mixture (linear superposition) of two-dimensional
Gaussians (MoGs).  MoGs are fast to render and fast to
affine-transform.  Most importantly, if you have a MoG PSF model
(which we advocate), the PSF-convolved, affine-transformed galaxy
models are themselves MoGs and therefore very fast to compute,
integrate, and render precisely.  We present worked examples that can
be directly used in image fitting; we are using them ourselves.  The
MoG profiles we provide can be swapped in to replace the standard
models in any image-fitting code; they sped up model fitting in our
projects by an order of magnitude; they ought to make any code faster
at very little cost in accuracy.
\end{abstract}

@@

The speed-ups that can be obtained by using MoG approximations can be
very large.  In our image-modeling project \project{The Tractor}, we
were PSF-convolving by rendering the profiles (especially the profile
centers) at very high resolution (hundreds to thousands of resolution
elements in the central pixel are necessary for good precision on the
dev profile).  We were then convolving that high-resolution model with
a low-resolution PSF and rendering to a low-resolution image pixel
grid.  These expensive operations were obviated by the MoG profiles,
which involve only rendering a small number of Gaussians at the pixel
centers on the low-resolution pixel grid.  The MoG approximations
saved us more than an order of magnitude in compute time, especially
in optimization, where derivatives have to be taken with respect to
the unconvolved model properties.
